Copulas and their potential for ecology

### 2019-10-03 Paper.Rmd
working on Q2 stuff

The first mechanism relates to the ideas in the Introduction about 
Liebig's law of the minimum, and to nonlinear influences of environmental
variables on ecological variables.
If an environmental variables influences an ecological variable 
disproportionately strongly in one of its tails, we explored whether the ecological
variable could then exhibit asymmetric tail dependence across space.
Let $E_i(t)$ be an ecological variable 
measured at location $i$ ($i=1,2$) at time $t$.
Assume the dynamics 
$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
the $\delta_i(t)$ are standard-normally distributed 
and independent across time and locations, $a=0.2$,
$b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
$\cov(\epsilon_1,\epsilon_2)=0.8$. Thus basic ecological dynamics follow the
very simple AR(1) formulation, influenced by a "regional" environmental factor which
is correlated across locatons ($\epsilon$) and by a local factor ($\delta$). 
The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
$E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
$g_1$ represents environmental effects that negatively impact populations,
and that occur only below a threshold of $\epsilon=0$; 
and $g_2$ represents effects that positively impact populations and 
that occur only above a threshold of $\epsilon=0$.
The values of $b$ selected provide a modicum of exploration of the potential
for ecological dynamics to influence how complex dependencies between
the $E_i(t)$ may arise: 
negative values of $b$ correspond to overcompensating dynamics 
and positive values to undercompensating dynamics; larger $|b|$
means slower return to equilibrium after a disturbance.

For each $b$ and $g_i$, we simulated the model for $25000$ time
steps and retained the $E_i(t)$ for the final $2500$ time 
steps. We applied our 
nonparametric statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
and our Spearman- or Kendall-preserving normal-copula 
surrogates (see section on methods for Q1) to these 
outputs to discover if the model could
produce asymmetric tail associations (and therefore non-normal copula structure)
between the $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
and $(\delta_1,\delta_2)$ have normal copula structure (they were drawn
from bivariate normal distributions), the Moran 
mechanism analyzed below does not operate here.
<!--***DAN: You may also want to have some additional commentary in the Discussion on
how the models and related models can be explored more comprehensively. In fact you may already have some of that, though you may want to beef it up. You may
want to insert a point at the end of the Methids section here again emphasizing
that these are simple models for an initial exploration, and we discuss in the 
Discussion how...-->

<!--***DAN: stopped here in Q2 methods in order to go on to the correspondng
section of Q2 results-->

Our second mechanism is an extension of the well known Moran
effect, and was summarized conceptually in the Introduction. 
Let $E_i(t)$ again be an ecological variable, $i=1,2$.
We initially assume the AR(1) dynamics 
$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
The environmental noises $\epsilon_i(t)$ were 
standard-normal random variables that were independent for distinct times, $t$, but 
exhibited different kinds of 
dependence across locations in different simulations (see below).
The variable $E$ is very general, and could represent deviations
of a population density from a carrying capacity, deviations of
total plant community biomass from an average value,
flux of a biogeochemical variable such as methane, or other quantities.

To explore the importance of nonlinear dynamics for our proposed mechanism, 
we also considered a
nonlinear population model $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$, using $r=0.5$, $K=100$, and $\sigma=0.1$ or 
$\sigma=1$. When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
$K$, and the nonlinearities of the model have limited influence on dynamics. When
$\sigma=1$, model dynamics are strongly nonlinear. We refer to these as
the weak-noise and strong-noise cases, though the importance of the noise here
is that, when it is strong, it brings the nonlinearities of the model into play. 
<!--***DAN: Didn't we use another value of r as well, and present results
on it? If so, probably should mention it here, in part so the exploration
seems a bit broader? Well, we mentioned that we did those and we verbally summarize
results but we did not show any. So maybe leave things as is in this respect.-->

For each of the model setups above, for each of several copula families, 
<!--***DAN: I currently use C, SC, G, SG, J, SJ. Do I really need all those? 
Surely it would be enough to just do C and SC and then state that the others 
give similar results? And if I move one of the result figure from this section
into the main text, that will mean only one such fig in the sup mat, at least
for AR(1)? I mean, right now we have 18 figure in sup mat from this analysis,
and only part of one of them in main text! See if the results are actually similar across
S/SC, G/SG, J/SJ, and if so consider keeping only the S/SC results in either the
main text or the sup mat, and just stating the others were similar. Then you would presumably
move the full C fig into the main text, replacing the current partial version.
So instead of having 18 figs in sup mat and half a fig in main text, you would
have 5 figs in sup mat and 1 in main text, and then you would just say "we did the
same thing for G, SG, J, SJ and got similar results."
So let's check, are results, broadly speaking, independent of which of the 6 copulas
I used?:
A) for AR(1) models, yes
B) for Ricker, weak noise, yes
C) for Ricker, strong noise, its more complicated. It's generally true that you get
more approximate agreement between noise and pop stats, with some rather glaring disagreement.
But which stats disagree the most depends on which cop we used!
So I guess we can still just display the C and SC results, as planned above, and explain
in text.
-->
for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
$50$ replicate simulations, we generated 5000 noise pairs 
$(\epsilon_1(t),\epsilon_2(t))$ from the
bivariate random variable with the given copula family and the given Kendall
correlation $\tau$.
We then used this noise to drive the model, and retained both the
noise and population values for the final 500 
time steps. For each simulation, the following
statistics were then computed for both noises and populations: 
Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
$\Dsq_l$, $\Dsq_u$, $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$. 
Values were plotted against $\tau$ for noises and populations. If the 
hypothesis from the Introduction was correct that characteristics of 
the copula structure of spatial dependence in an
ecological variable can be inherited from the copula structure of spatial 
dependence in an environmental variables, then 
plots should be similar for populations and noises.

@@

Our model with asymmetric environmental effects 
produced outputs with visually apparent asymmetry 
of tail associations between the ecological variables $E_i$ in the two locations, 
to an extent that depended on the value of $b$.
For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
for both $g_1$ and $g_2$,
asymmetry of tail association was strong; for 
$b=\pm 0.5$, asymmetry was weaker but still apparent
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
Lower-tail (respectively, upper-tail) spatial associations in the 
effects of noise ($g_1$, respectively, $g_2$) 
produced lower-tail (respectively, upper-tail) spatial 
associations in the ecological variable, $E_i$.
Results using our statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
strongly reflected the visually apparent asymmetry (Tables \ref{tab_stat_npa_asym}). 
Thus asymmetry of environmental effects is a mechanism that may be partly 
responsible for non-normal copula structure and
asymmetric tail dependence across space in ecological variables.
This result is represented in Fig. \ref{MasterFigure} as the solid box around 
"Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
It is explained in the Discussion why the box and some of those arrows 
are solid, instead of dashed, although the results we 
present so far in this section are theoretical.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
\caption{If environmental effects operate asymmetrically in their tails on
ecological variables, it can result in non-normal copula structure and asymmetric 
tail dependence across space in the ecological variables. Shown are the last
$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$.\label{fig_ms_asym_b_0.1}}
\end{center}
\end{figure}

@@

<!--***DAN: got to here so far in editing the result for Q2-->

<!--Second mechanism, the Moran one-->
The hypothesis that characteristics of 
spatial copula structure and tail dependence
in an ecological variable can be inherited from an environmental 
variable was largely confirmed; 
<!--***DAN: You are using language about confirming hypotheses, here and elsewhere,
but this may rub some people the wrong way because we have not confirmed anything
empirically, we are just doing models, and only a cursory exploration of each
model, at that! So change the language.-->
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
Thus this is a mechanism that can produce non-normal copula structure and
asymmetric tail dependence 
<!--****DAN: Don't firget to replace @tail dependence@ by @tail asociation@ as appropriate-->
across space in ecological variables.
<!--***DAN: move the below to end of this part-->
This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
"Moran effects" and the arrows labelled "B".
<!--***DAN: move the above to end of this part-->
For simulations using AR(1) models, copula 
statistics were always very similar for noise and populations 
(Fig \ref{fig_Cause4copula_scatter_C_kend_ar1}, 
\ref{SM-fig_Cause4copula_C_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1}),
though there were significant but minor differences for some statistics
and simulations.
For weak noise and our nonlinear model, copula statistics
were again very similar for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}).
Though there were again significant but small differences for some statistics
and simulations, the hypothesis was supported that copula 
structure in the noise translates
to similar but not identical copula structure in output populations. 
Since many ecological models are nonlinear, this result provides the 
expectation that there will be a Moran-effect-like correspondence 
between noise and model-output copula structure across space 
for typical ecological dynamics,
as long as noise is small enough that dynamics stay relatively close
to the model equilibrium. Results that hold for "weak noise" in this sense
are common in ecology. 
We repeated the analysis with $r=1.3$, with similar results. The deterministic 
one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
dynamics, e.g., $r=1.3$).

\begin{figure}[!h]
\begin{center}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption{Example results showing similarity of copula statistics for 
environmental-noise inputs and ecological-variable outputs of a dynamical model.
Asymmetric tail dependence was similar in both noise inputs and model outputs
for this model. The 
AR(1) model and a Clayton copula were used (see methods section for Q2).
Each point is the mean across 50 replicate simulations for which the same 
Kendall $\tau$ value was used. Error bars are standard errors 
and panel headers give Pearson correlation results for
the points. The regression line through the points (black line) was similar
to the 1-1 line (green line). 
See Fig. \ref{SM-fig_Cause4copula_C_kend_ar1} for additional results for the AR(1) model and Clayton copula, 
Figs \ref{SM-fig_Cause4copula_SC_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1} for AR(1) results with other copula families, and 
Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
the nonlinear models described in the methods section for Q2.\label{fig_Cause4copula_scatter_C_kend_ar1}}
\end{center}
\end{figure}

For strong noise and using our nonlinear model, 
copula statistics, generally speaking, were approximately similar 
between noise and model
outputs; however, similarity was reduced relative to previous results, 
and for a few simulations, tail dependence asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). For instance, using a Clayton  copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}); 
and likewise for survival Gumbel and survival Joe copulas (Figs \ref{SM-fig_Cause4copula_SG_kend_sRicker_nsd_1}, 
\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1}). Overall,
similarities between noise and model-output copula statistics
were more noticable than discrepancies, in spite of nonlinearities.
We repeated the analysis with $r=1.3$. Discrepancies between noise and population
copula structure in this case were often glaring. Apparently 
noise of standard deviation 1 interacted especially strongly with 
model nonlinearities when the model was in its overcompensatory 
regime.

>>>>

The first mechanism relates to the ideas in the Introduction about 
Liebig's law of the minimum, and to nonlinear influences of environmental
variables on ecological variables.
If an environmental variables influences an ecological variable 
disproportionately strongly in one of its tails, we explored whether the ecological
variable could then exhibit asymmetric tail dependence across space.
Let $E_i(t)$ be an ecological variable 
measured at location $i$ ($i=1,2$) at time $t$.
Assume the dynamics 
$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
the $\delta_i(t)$ are standard-normally distributed 
and independent across time and locations, $a=0.2$,
$b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
$\cov(\epsilon_1,\epsilon_2)=0.8$. Thus basic ecological dynamics follow the
very simple AR(1) formulation, influenced by a "regional" environmental factor which
is correlated across locatons ($\epsilon$) and by an entirely local factor ($\delta$). 
The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
$E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
$g_1$ represents environmental effects that negatively impact populations,
and that occur only below a threshold of $\epsilon=0$; 
and $g_2$ represents effects that positively impact populations and 
that occur only above a threshold of $\epsilon=0$.
The values of $b$ selected provide a modicum of exploration of the potential
for ecological dynamics to influence how complex dependencies between
the $E_i(t)$ may arise: 
negative values of $b$ correspond to overcompensating dynamics 
and positive values to undercompensating dynamics; larger $|b|$
means slower return to equilibrium after a disturbance.

For each $b$ and $g_i$, we simulated the model for $25000$ time
steps and retained the $E_i(t)$ for the final $2500$ time 
steps. We applied our 
nonparametric statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
and our Spearman- or Kendall-preserving normal-copula 
surrogate comparison methods (see section on methods for Q1) to these 
outputs to discover if the model could
produce asymmetric tail associations (and therefore non-normal copula structure)
between the $E_i(t)$. Because $(\epsilon_1,\epsilon_2)$
and $(\delta_1,\delta_2)$ have normal copula structure (they were drawn
from bivariate normal distributions), the Moran 
mechanism analyzed below does not operate here.
<!--***DAN: You may also want to have some additional commentary in the Discussion on
how the models and related models can be explored more comprehensively. In fact you may already have some of that, though you may want to beef it up. You may
want to insert a point at the end of the Methids section here again emphasizing
that these are simple models for an initial exploration, and we discuss in the 
Discussion how...-->

Our second mechanism is an extension of the well known Moran
effect, and was summarized conceptually in the Introduction. 
We consider a linear model, as well as 
two parameterizations of a nonlinear population model which includes density dependence
of population growth. For the linear model, let $E_i(t)$ 
again be an ecological variable, $i=1,2$.
We use AR(1) dynamics, 
$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
The environmental noises $\epsilon_i(t)$ were 
standard-normal random variables that were independent for distinct times, $t$, but 
exhibited different kinds of 
dependence across locations in different simulations (see below).
The variable $E$ is very general, and could represent deviations
of a population density from a carrying capacity, deviations of
total plant community biomass from an average value,
flux of a biogeochemical variable such as methane, or other quantities.
The nonlinear population model was $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$ for $i=1,2$, using $r=0.5$, $K=100$, $\epsilon_i(t)$
as above, and $\sigma=0.1$ or 
$\sigma=1$ in different simulations. 
When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
$K$, where the model equation can be well approximated by a linear equation,
and the nonlinearities of the model therefore have limited influence on dynamics. When
$\sigma=1$, model dynamics are strongly nonlinear because the stochastic 
component of the model causes populations to stray far from the carrying capacity, $K$. 
We refer to these as
the weak-noise and strong-noise cases, though the importance of the noise here
is that, when it is strong, it brings the nonlinearities of the model into play.

For each of the three model setups above (the linear model and the nonlinear model
with weak and strong stochasticity), for the Clayton and survival Clayton copula families, 
for each $\tau=0.1,0.2,\ldots,0.9$, and for each of
$50$ replicate simulations, we generated 5000 noise pairs 
$(\epsilon_1(t),\epsilon_2(t))$ from the
bivariate random variable with standard-normal marginals and with 
the given copula family and the given Kendall
correlation $\tau$.
We then used this noise to drive the model, and retained both the
noise and population values for the final 500 
time steps. For each simulation, the following
statistics were then computed for both noises and populations: 
Pearson, Spearman and Kendall correlations, $\cor_l$, $\cor_u$, $\Ps_l$, $\Ps_u$, 
$\Dsq_l$, $\Dsq_u$, $\cor_l - \cor_u$, $\Ps_l - \Ps_u$, and $\Dsq_u - \Dsq_l$. 
Values were plotted against $\tau$ for noises and populations. If the 
hypothesis from the Introduction was reasonable that characteristics of 
the copula structure of spatial dependence in an
ecological variable may be inherited from characteristics of spatial 
dependence in an environmental variable through a Moran-like effect, then 
plots should be similar for populations and noises.

<!--***DAN: stopped here for methods of Q2-->

@@

Our model with asymmetric environmental effects 
produced outputs with visually apparent asymmetry 
of tail associations between the ecological variables $E_i$ in the two locations, 
to an extent that depended on the value of $b$.
For $b=0.1$ (Fig. \ref{fig_ms_asym_b_0.1}) and $b=-0.1$ 
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}A,B), 
for both $g_1$ and $g_2$,
asymmetry of tail association was strong; for 
$b=\pm 0.5$, asymmetry was weaker but still apparent
(Fig. \ref{SM-fig_ms_asym_b_-0.1_0.5_-0.5}C-F).
Lower-tail (respectively, upper-tail) spatial associations in the 
effects of noise ($g_1$, respectively, $g_2$) 
produced lower-tail (respectively, upper-tail) spatial 
associations in the ecological variable, $E_i$.
Results using our statistic $\cor_{0,0.1}-\cor_{0.9,1}$ 
strongly reflected the visually apparent asymmetry (Table \ref{tab_stat_npa_asym}). 
Thus asymmetry of environmental effects is a mechanism that may be partly 
responsible for non-normal copula structure and
asymmetric tail associations across space in ecological variables.
This result is represented in Fig. \ref{MasterFigure} as the solid box around 
"Nonlinear environmental effects, Liebig's law" and the arrows labelled "A".
It is explained in the Discussion why the box and some of the arrows 
are solid, instead of dashed, although the results of this section are theoretical.

\begin{figure}[!h]
\begin{center}
\includegraphics[width=12 cm]{./Results/asym_sens_results/a_0.2_b_0.1_r_0.8_asym_sens.pdf}
\caption{If environmental effects operate asymmetrically in their tails on
ecological variables, it can result in non-normal copula structure and asymmetric 
tail dependence across space in the ecological variables. Shown are the last
$500$ points for (A) $g=g_1$ and (B) $g=g_2$ from simulations described in the text, $b=0.1$. 
Asymmetric tail associations are visually apparent, but were also elaborated statistically in 
Table \ref{tab_stat_npa_asym}. \label{fig_ms_asym_b_0.1}}
\end{center}
\end{figure}

@@

<!--Second mechanism, the Moran one-->
The hypothesis that characteristics of 
the copula structure of spatial dependence 
in an ecological variable can be inherited from an environmental 
variable was found to be reasonable, because it held for the models we examined; 
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
For simulations using AR(1) models, our correlation and tail asymmetry statistics 
were always similar for noise and populations 
(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayon copula, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
<!--***DAN: reorder sup mat figs so they are cited in order-->
Though there were significant differences for many statistics
and simulations, these were small compared to the overall tendency for
larger (respectively, smaller) values of our statistics, as applied to
noise, to be paired with larger (respectively, smaller) values of the same
statistics as applied to model outputs. 
For our nonlinear model with weak noise, values of the statistics
were again quite similar for noise and model outputs 
(Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1} for Clayton, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_0.1} for survival Clayton).
Though there were again significant differences for many statistics
and simulations, these were again small relative to overall variation of 
values of the statistics.
Since many ecological models are nonlinear, this result provides the 
reasonable expectation that there may be a Moran-effect-like correspondence 
between noise and model-output dependence structure across space 
for typical ecological dynamics,
as long as environmental noise is small enough that dynamics stay relatively close
to the model equilibrium. Results that hold for "weak noise" in this sense
are common in ecology.
For strong noise and using our nonlinear model, 
our correlation and tail-asymmetry statistics, generally 
speaking, were approximately similar between noise and model
outputs; however, similarity was reduced relative to previous results, 
and for a few simulations, asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1} and \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_1}). For instance, 
using a Clayton  copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).

We repeated our analyses using the nonlinear model with $r=1.3$. The deterministic 
one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
dynamics, e.g., $r=1.3$). For weak noise, similarities were again dominant between
values of our statistics on noise and population time series. 
For strong noise, however, discrepancies were often glaring. Apparently 
noise of standard deviation 1 interacted especially strongly with 
model nonlinearities when the model was in its overcompensatory 
regime. We repeated all analyses using the Gumbel, survival Gumbel, Joe, and 
survival Joe copulas, with substantially similar main conclusions (not shown).

Thus it is reasonable to hypothesize that a Moran-effect-like 
mechanism may produce non-normal copula structure and
asymmetric tail associations across space in ecological variables.
This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
"Moran effects" and the arrows labelled "B".

<!--***DAN: Decided to remove this simplified version of the figure for AER
and replace by the bigger one which was in the supp mat, now below.
\begin{figure}[!h]
\begin{center}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption{Example results showing similarity of copula statistics for 
environmental-noise inputs and ecological-variable outputs of a dynamical model.
Asymmetric tail dependence was similar in both noise inputs and model outputs
for this model. The 
AR(1) model and a Clayton copula were used (see methods section for Q2).
Each point is the mean across 50 replicate simulations for which the same 
Kendall $\tau$ value was used. Error bars are standard errors 
and panel headers give Pearson correlation results for
the points. The regression line through the points (black line) was similar
to the 1-1 line (green line). 
See Fig. \ref{SM-fig_Cause4copula_C_kend_ar1} for additional results for the AR(1) model and Clayton copula, 
Figs \ref{SM-fig_Cause4copula_SC_kend_ar1}-\ref{SM-fig_Cause4copula_SJ_kend_ar1} for AR(1) results with other copula families, and 
Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1}-\ref{SM-fig_Cause4copula_SJ_kend_sRicker_nsd_1} for
the nonlinear models described in the methods section for Q2.\label{fig_Cause4copula_scatter_C_kend_ar1}}
\end{center}
\end{figure}
-->

<!--Plots for AR(1) with Clayton-->
\begin{figure}[!h]
\begin{center}
\includegraphics[width=13 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/common_legend_cause4copula_stat.pdf}\\
\vspace{-0.4 cm}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Spearman_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Kendall_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pearson_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Coru_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Corl-Coru_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pu_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_Pl-Pu_vs_Kendall's Tau.pdf}\\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2l_vs_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_D2u-D2l_vs_Kendall's Tau.pdf} \\
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_CorlmCoru_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_PlmPu_Kendall's Tau.pdf}
\includegraphics[width=5 cm]{./Results/Cause4copula_results/Cause4copula_stat_results_ar1_beta_0.5_Clayton/Clayton_scatter_D2umD2l_Kendall's Tau.pdf}
\caption{Comparison of correlation and tail association statistics between 
environmental-noise inputs and ecological-variable 
outputs, AR(1) model, Clayton copula. Red and blue points give 
means over 50 replicate simulations of the listed statistics, 
red points are for noise inputs and blue points are for
model outputs. Error bars are standard errors. 
$P$-values (purple points, right axis) are for a 
paired $t$-test of the null hypothesis that the distributions have the same mean. 
For bottom panels, headers give Pearson correlations for
the points. The regression line through the points (black) was similar
to the 1-1 line (green). 
See methods section for Q2 
for details.\label{fig_Cause4copula_C_kend_ar1}}
\end{center}
\end{figure}
<!--***DAN: Shyamolina, the panels in the bottom row do not line up with the others. 
I remember you fixed this before for the sup mat figs, can you plase fix it again here?
Also, can you make the panels smaller or more tightly packed or soemthing
so the whole figure plus caption fits on a page?-->

<!--***DAN: Got to here in results for Q2-->
